abxD01 Paper

### 2015-03-07 PAPER.Rmd
Edit titration analysis

***Statistical Analysis & Modeling*** Random forest regression models were constructed using the randomForest R package using XX,000 trees [ref]. The
model was constructed using the relative abundance of those OTUs that had an
average relative abundance within any antibiotic treatment group of 1%. This
resulted in 44 OTUs. The regression was performed using the log10 transformation
of the number of CFU per gram of fecal material as the dependent variable. All analyses were conducted using R version 3.1.2. Complete analysis scripts are
available at the online repository for this study (https://github.com/SchlossLab/Schubert_AbxD01_mBio_2015).

@@

***Smaller perturbations of the microbiota result in maintenance of colonization resistance.*** <transition>
Because most of our original treatments resulted in similarly high levels of *C. difficile*, we hypothesized that applying smaller antibiotic perturbations to the microbiota would result in reduced sensitivity to *C. difficile*, as we observed in the ciprofloxacin treated mice. To test this, we titrated the doses of cefoperazone, streptomycin, and vancomycin given to mice (Supplemental Table 1). Colonization levels decreased significantly in all mice receiving titrated doses in support of our hypothesis. *C. difficile* levels were significantly different between all titration groups within the cefoperazone (p<0.03) and streptomycin treated microbiotas (p<0.01), displaying a greater range of colonization resistance profiles (Figure 3). Next, we hypothesized that the populations we previously identified as being associated with protection against colonization would be elevated in the mice with decreased colonization and those associated with susceptibility would be less abundant.

The titrated antibiotic doses clearly resulted in distinct community structures that were variations of the original dose (Figure 3). The cefoperazone titrations resulted in a shift in dominance from the *Bacteroidetes* to the *Firmicutes* phylum with increasing antibiotic dose. Most notable was the increase in OTUs associated with resistance including the *Porphyromonadaceae* and *Alistipes* and a reduction in OTUs associated with susceptibility including members of the *Proteobacteria*, *Clostridium*, and *Staphylococcus*. The relative abundance of other OTUs associated with protection including the *Lactobacillus*, *Lachnospiraceae*, and *Ruminococcaceae* did not change as expected based on the initial correlation analysis; however, these OTUs were present at levels similar to the untreated mice. The streptomycin titrations did not result in profound shifts in the overall community structure, but did yield increased resistance. Rather, there were individual populations that were gained or lost across the gradient. Notably, mice receiving the lowest dose retained significantly higher levels of *Alistipes* and *Anaeroplasma* (p<0.05), which we previously found to be strongly associated with colonization resistance. The relative abundances of the *Porphyromonodaceae*, *Lachnospiraceae*, *Ruminococcaceae* were unchanged, but were present at similarly high levels to those found in the untreated mice. Although the vancomycin titration yielded a number of large changes in the abundance of individual populations, there was no recovery of colonization resistance. The relative abundance of many of the populations associated with resistance (*Porphyromonadaceae*, *Lachnospiraceae*, *Anaeroplasma*) increased with decreased vancomycin dosage; however, with the exception of the *Lactobacillus* populations, they did not reach the levels observed in the untreated mice. In addition, all of the vancomycin-treated mice had high levels of OTUs affiliated with the *Akkermansia* and *Enterobacteriaceae*. Together, these results suggest that individual populations are not sufficient to provide colonization resistance. Rather, resistance is likely a product of the overall composition of the community.

@@

**Figure 3. TITLE.** This figure shows the average percent relative abundances for family level phylotypes found on the day of *C. difficile* challenge in each titration group. Each row depicts a different antibiotic that was titrated. High, medium, and low antibiotic doses are represented by black, grey, and white colored bars respectively. The specific dose amount is shown for each antibiotic. Error bars show the standard error of the mean.

>>>>

***Statistical Analysis & Modeling*** All analyses were conducted using R version 3.1.2. Complete analysis scripts are available at the online repository for this study (https://github.com/SchlossLab/Schubert_AbxD01_mBio_2015).

Need text about...
* correlation analysis
* comparison of titration levels
* For the random forest you use 1% within each treatment group, for the correlation you state using 0.1 and 0.01% across all samples. Need to stay consistent.
* statistical methods used to differentiate the abundances of OTUs in the titrations

Random forest regression models were constructed using the randomForest R package using XX,000 trees [ref]. The model was constructed using the relative abundance of those OTUs that had an average relative abundance within any antibiotic treatment group of 1%. This resulted in 44 OTUs. The regression was performed using the log10 transformation of the number of CFU per gram of fecal material as the dependent variable.

@@

***Smaller perturbations of the microbiota result in maintenance of colonization resistance.*** All but the ciprofloxacin treatment resulted in similarly high levels of *C. difficile*. Therefore, we hypothesized that titrating the dose of antibiotics given to the mice would result in a reduced perturbation to the microbiota and a reduced sensitivity to *C. difficile*. We titrated the doses of cefoperazone, streptomycin, and vancomycin given to the mice (Supplemental Table 1). Colonization levels decreased significantly in all mice receiving titrated doses in support of our hypothesis (Figure 3). *C. difficile* levels were significantly different between all titration groups within the cefoperazone (p<XXX0.03XXX) and streptomycin treated mice (p<XXX0.01XXX). For each of the antibiotics we titrated, reduced dosages resulted in reduced colonization by *C. difficile*. By sequencing the 16S rRNA genes from feces collected on day 0 from these mice, we sought to validate the relationships we observed with the full dosage. Titrating cefoperazone resulted in a shift in dominance from the *Bacteroidetes* to the *Firmicutes* phylum with increasing antibiotic dose. Most notable was the increase in the relative abundance of OTUs associated with resistance including the *Porphyromonadaceae* and *Alistipes* and a reduction in OTUs associated with susceptibility including members of the *Proteobacteria*, *Clostridium*, and *Staphylococcus*. The relative abundance of other OTUs associated with protection including the *Lactobacillus*, *Lachnospiraceae*, and *Ruminococcaceae* did not change as expected based on the initial correlation analysis; however, these OTUs were present at levels similar to the untreated mice. The at the lowest streptomycin titration, mice retained significantly higher abundances of *Alistipes* and *Anaeroplasma* (p<0.05), which we found to be strongly associated with colonization resistance. The relative abundances of the *Porphyromonodaceae*, *Lachnospiraceae*, *Ruminococcaceae* were unchanged, but were present at similarly high levels to those found in the untreated mice. Although the vancomycin titration yielded a number of large changes in the abundance of individual populations, there was no recovery of colonization resistance. The relative abundance of many of the populations associated with resistance (*Porphyromonadaceae*, *Lachnospiraceae*, *Anaeroplasma*) increased with decreased vancomycin dosage; however, with the exception of the *Lactobacillus* populations, they did not reach the levels observed in the untreated mice. In addition, all of the vancomycin-treated mice had high levels of OTUs affiliated with the *Akkermansia* and *Enterobacteriaceae*. Together, these results suggest that individual populations are not sufficient to provide colonization resistance. Rather, resistance is likely a product of the overall composition of the community.

@@

**Figure 3. The average percent relative abundances for family level phylotypes found on the day of *C. difficile* challenge in each titration group.** Each row depicts a different antibiotic that was titrated. High, medium, and low antibiotic doses are represented by black, grey, and white colored bars respectively. The specific dose amount is shown for each antibiotic. Error bars show the standard error of the mean.
